\stktokens{}: Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities

### 2018-10-17 fully-abstract-comp/popl2019/paper.tex
phase-2 version should no longer be anonymous
More work on figure 10.

%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%\documentclass[acmsmall]{acmart}\settopmatter{}
%% For editing double-blind review submission, w/o CCS and ACM Reference (max submission space)
\documentclass[acmsmall,review,anonymous,showframe]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
%\documentclass[acmsmall]{acmart}\settopmatter{}

@@

\begin{figure}[htb]
  \centering
  \begin{mathpar}
    \inferrule{ \src{\Phi(\pcreg) = ((\perm,\_),\baddr,\eaddr,\aaddr)} \\
               \src{\lbrack \aaddr ,\aaddr + \calllen - 1\rbrack \subseteq \ta}  \\
               \src{\lbrack\aaddr,\aaddr + \calllen-1\rbrack \subseteq [\baddr,\eaddr]} \\
               \src{\perm \in \{\rwx,\rx\}} \\
               \src{\Phi.\mem(\aaddr,\dots,\addr + \calllen-1) = \scall[0]{\offpc,\offsigma}{r_1}{r_2} \cdots \scall[\calllen-1]{\offpc,\offsigma}{r_1}{r_2}}
 }
              { \src{\Phi}  \; \src{\step[\src{\ta,\stkb}] \sem{\scall{\offpc,\offsigma}{r_1}{r_2}}(\Phi)} }
  \end{mathpar}
  \begin{tabular}{|>{$}c<{$}|p{3.7cm}|>{\raggedright\arraybackslash}p{6.55cm}|}
    \hline
    i \in \Instr                                 & $\sem{i}(\Phi)$ & Conditions\\
    \hline
    \halt                                        & $\halted$ & \\
    \hline
    \multicolumn{3}{|c|}{\dots\text{ (the operational semantics of \trgcm)}} \\
    \hline
    \store{r_1}{r_2}                             & $\srcalt{\var{updPc}} \sourcecolor\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_2}{w_2}\\ &\update{\ms_\stk.\aaddr}{\Phi(r_2)})\endarray$  & \srcalt{$\Phi(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $\perm \in \{\rwx,\rw\}$ and $\baddr \le \aaddr \le \eaddr$ and $w_2 = \linCons{\Phi(r_2)}$ and $\aaddr \in \dom(\ms_\stk)$}\\
    \hline
    \cca{r}{\rn}                                 &$\srcalt{\updPcAddr{\Phi\updReg{r}{w}}}$ &  \srcalt{$\Phi(\rn) = n \in \ints$ and $\Phi(r) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $w = \stkptr{\perm,\baddr,\eaddr,\aaddr+n}$} \\
    \hline
    \scall{\srcalt{\offpc,\offsigma}}{\src{r_1}}{\src{r_2}} &
        $\srcalt{\var{xjmpRes}(c_1,c_2,}$
        $\srcalt{\hspace{.25cm}\left(\arraycolsep=0pt\array[c]{rl}\Phi&\updReg{r_1,r_2}{w_1,w_2}\\
            &\updReg{\rretc}{s_c}\\
            &\updReg{\rretd}{s_d}\\
            &\updReg{\rstk}{c_\stk}\\
            &\update{\ms_\stk}{\ms_{\stk,\var{rest}}}\\
            &\update{\stk}{\stk'}\\
            \endarray\right))}$
       &
      \srcalt{$\stk' = (\opc,\ms_{\stk,\var{local}}) :: \Phi.\stk$ and 
         $\begin{multlined}
           \ms_{\stk,\var{local}},\ms_{\stk,\var{rest}},c_\stk,c_\var{local} =\\ \var{splitStack}(\Phi.\reg(\rstk), \Phi.\ms_\stk) \text{ and }
         \end{multlined}$
    $\opc, c_\opc =\var{setupOpc(\Phi.\reg(\pcreg))}$ and
    $\sigma = \var{getCallSeal}(\Phi.\reg,\Phi.\mem,\offpc,\offsigma)$ and 
    $s_c,s_d = \var{sealReturnPair}(\sigma,c_\var{local},c_\opc)$
    $w_1,w_2 = \linCons{\Phi.\reg(r_1,r_2)}$ and
                        $\reg(r_1,r_2) = \sealed{\sigma', c_1}, \sealed{\sigma', c_2}$
                        }\\
    \hline 
    \multicolumn{3}{|c|}{\dots} \\
    \hline
    \_                                           & $\failed$ & \totherwise \\
    \hline
  \end{tabular}
\begin{multline*}
  \srcxjmpres{c_1,c_2,\Phi} = \\
  \left\{ 
    \arraycolsep=1.4pt
    \begin{array}{l c >{\raggedright\arraybackslash}p{9.5cm}}
      \arraycolsep=0pt
    \begin{array}{rl}
      \Phi & \updReg{\pcreg}{c_1}\\
           & \updReg{\rdata}{c_2}
    \end{array}
    &\phantom{mak} &  $\nonExec{c_2}$ \srcalt{and $c_1 \neq \retptrc(\_)$ and $c_2 \neq \retptrd(\_)$}\\
        \src{\Phi'} 
      & &
        $\src{ \Phi' = \var{returnCondition}(c_1,c_2,\Phi)}$ \\
      \failed &  & \totherwise
    \end{array} 
\right.
\end{multline*}
  \caption{An excerpt of the operational semantics of \srccm{} (some details omitted).}
  \label{fig:source-op-sem}
\end{figure}

\begin{figure}
    \[
      \src{\var{sealedPairArgument}(\reg,r_1,r_2,c_1,c_2)} \mathit{ iff }
      \left\{ 
      \begin{array}{l}
        \src{\exists \sigma \ldotp (\reg(r_1) = \sealed{\sigma, c_1} \wedge \reg(r_2) = \sealed{\sigma, c_2}) \wedge{}}\\
        \src{r_1 \neq \rtmp{1} \wedge r_2 \neq \rtmp{2}}
      \end{array}
      \right.
    \]
    \[
      \src{ \opc, c_\opc = \var{setupOpc(((\_,\_),\baddr,\eaddr,\aaddr))}} \mathit{ iff }
      \left\{ 
      \begin{array}{l}
      \src{\opc = \aaddr + \calllen \wedge{}}\\
      \src{c_\opc = \retptrc(\baddr,\eaddr,\opc) \wedge{}}
      \end{array}
       \right.
    \]
    \begin{multline*}
      \src{\ms_{\stk,\var{local}},\ms_{\stk,\var{unused}},c_\stk,c_\var{local\_data} = \var{splitStack}(\stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk}, \ms_\stk)} \mathit{\ iff} \\
      \left\{ 
      \begin{array}{l}
      \src{\baddr_\stk < \aaddr_\stk \leq \eaddr_\stk }\\
      \src{\ms_{\stk,\var{local}} = \ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42} }\\
      \src{\ms_{\stk,\var{unused}} = \ms_\stk|_{[\baddr_\stk,\aaddr_\stk-1]} }\\
      \src{c_\stk = \stkptr{\rw,\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1} }\\
      \src{c_\var{local\_data} = \retptrd(\aaddr_\stk,\eaddr_\stk) }\\
      \end{array}
      \right.
    \end{multline*}
    \[
      \src{\sigma = \var{getCallSeal}(\reg,\mem,\offpc,\offsigma)} \mathit{\ iff\ }
      \left\{
      \begin{array}{l} 
      \src{\reg(\pcreg) = ((\_,\_),\baddr,\eaddr,\aaddr) \wedge \baddr \leq  \aaddr+\offpc \leq \eaddr \wedge{}} \\
      \src{\mem(\aaddr+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\aaddr} \wedge{} \sigma_\baddr \leq \sigma \leq \sigma_\eaddr \wedge{}}\\
      \src{\sigma = \sigma_\aaddr + \offsigma }
      \end{array}
      \right.
    \]

@@

  \caption{The call condition in the operational semantics of \srccm{}.}
  \label{fig:source-op-sem}
\end{figure}
\begin{figure}
\[
    \src{\stkptr{\rw,\stkb,\eaddr_\stk,\aaddr_\stk} = \var{tokenToStkPtr}(\retptrd(\aaddr_\stk,\eaddr_\stk)}
\]
\[
    \src{((\rx,\normal),\baddr,\eaddr,\opc) = \var{tokenToPc}(\retptrc(\baddr,\eaddr,\opc)}
\]
\[
    \src{\stk', \ms_\var{local\_stk} = \var{popCallStackFrame} (\stk,c_1,c_2)}
    \mathit{\ iff\ }
      \left\{ \array{l}
        \src{\stk = (\opc, \ms_\var{local\_stk}) :: \stk' \wedge{}} \\
        \src{c_1 = \retptrc(\_,\_,\opc) \wedge{}} \\
        \src{c_2 = \retptrd(\aaddr_\stk,\eaddr_\stk)\wedge{}} \\
        \src{\dom(\ms_\var{local\_stk}) \subseteq [\aaddr_\stk,\eaddr_\stk]}
        \endarray
     \right.
\]
\[
    \src{c_\stk = \var{reconstructStackPointer}(\reg,c)}
    \mathit{\ iff\ }
      \left\{ \array{l}
        \src{c = \retptrd(\aaddr_\stk,\eaddr_\stk) \wedge{}} \\
        \src{\reg(\rstk) = \stkptr{\rw, \stkb, \aaddr_\stk-1,\_} \wedge{}} \\
        \src{\stkb \leq \addr_\stk \wedge{}} \\
        \src{c_\stk = \var{tokenToStkPtr(c)} }
        \endarray
     \right.
\]

\begin{multline*}
    \src{(\mem,\reg',\stk',\ms_\stk') = \var{returnCondition}(c_1,c_2,(\mem,\reg,\stk,\ms_\stk))}\mathit{\ iff\ }\\
      \quad\quad\left\{ 
      \begin{array}{l}
      \src{\stk', \ms_\var{local\_stk} = \var{popCallStackFrame}(\stk,c_1,c_2)}\\
      \src{c_\stk = \var{reconstructStackPointer}(\reg,c_2)}\\
      \src{c_\opc =  \var{tokenToPc}(c_1)} \\
      \src{\ms_\stk' = \ms_\stk \uplus \ms_\var{local\_stk}} \\
      \src{\reg' =\reg\updReg{\pcreg,\rstk}{c_\opc,c_\stk}\updReg{\rdata,\rtmp{1},\rtmp{2}}{0,0,0}}
      \end{array}\right.
  \end{multline*}

>>>>

%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%\documentclass[acmsmall]{acmart}\settopmatter{}
%% For editing double-blind review submission, w/o CCS and ACM Reference (max submission space)
\documentclass[acmsmall,review,showframe]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
%\documentclass[acmsmall]{acmart}\settopmatter{}

@@

\begin{figure}[htb]
    \inferrule{ \src{\Phi(\pcreg) = ((\perm,\_),\baddr,\eaddr,\aaddr)} \\
               \src{\lbrack \aaddr ,\aaddr + \calllen - 1\rbrack \subseteq \ta}  \\
               \src{\lbrack\aaddr,\aaddr + \calllen-1\rbrack \subseteq [\baddr,\eaddr]} \\
               \src{\perm \in \{\rwx,\rx\}} \\
               \src{\Phi.\mem(\aaddr,\dots,\addr + \calllen-1) = \scall[0]{\offpc,\offsigma}{r_1}{r_2} \cdots \scall[\calllen-1]{\offpc,\offsigma}{r_1}{r_2}}
 }
              { \src{\Phi}  \; \src{\step[\src{\ta,\stkb}] \sem{\scall{\offpc,\offsigma}{r_1}{r_2}}(\Phi)} }
  \begin{tabular}{|>{$}c<{$}|p{3.7cm}|>{\raggedright\arraybackslash}p{6.55cm}|}
    i \in \Instr                                 & $\sem{i}(\Phi)$ & Conditions\\
    \halt                                        & $\halted$ & \\
    \multicolumn{3}{|c|}{\dots\text{ (the operational semantics of \trgcm)}} \\
    \store{r_1}{r_2}                             & $\srcalt{\var{updPc}} \sourcecolor\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_2}{w_2}\\ &\update{\ms_\stk.\aaddr}{\Phi(r_2)})\endarray$  & \srcalt{$\Phi(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $\perm \in \{\rwx,\rw\}$ and $\baddr \le \aaddr \le \eaddr$ and $w_2 = \linCons{\Phi(r_2)}$ and $\aaddr \in \dom(\ms_\stk)$}\\
    \cca{r}{\rn}                                 &$\srcalt{\updPcAddr{\Phi\updReg{r}{w}}}$ &  \srcalt{$\Phi(\rn) = n \in \ints$ and $\Phi(r) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $w = \stkptr{\perm,\baddr,\eaddr,\aaddr+n}$} \\
    \scall{\srcalt{\offpc,\offsigma}}{\src{r_1}}{\src{r_2}} &
        $\srcalt{\var{xjmpRes}(c_1,c_2,}$
        $\srcalt{\hspace{.25cm}\left(\arraycolsep=0pt\array[c]{rl}\Phi&\updReg{r_1,r_2}{w_1,w_2}\\
            &\updReg{\rretc}{s_c}\\
            &\updReg{\rretd}{s_d}\\
            &\updReg{\rstk}{c_\stk}\\
            &\update{\ms_\stk}{\ms_{\stk,\var{rest}}}\\
            &\update{\stk}{\stk'}\\
            \endarray\right))}$
       &
      \srcalt{$\stk' = (\opc,\ms_{\stk,\var{local}}) :: \Phi.\stk$ and 
         $\begin{multlined}
           \ms_{\stk,\var{local}},c_\var{local},\ms_{\stk,\var{rest}},c_\stk =\\ \var{splitStack}(\Phi.\reg(\rstk), \Phi.\ms_\stk) \text{ and }
         \end{multlined}$
    $\opc, c_\opc =\var{setupOpc(\Phi.\reg(\pcreg))}$ and
    $\sigma = \var{getCallSeal}(\Phi.\reg,\Phi.\mem,\offpc,\offsigma)$ and 
    $s_c,s_d = \var{sealReturnPair}(\sigma,c_\var{local},c_\opc)$ and
    $w_1,w_2 = \linCons{\Phi.\reg(r_1,r_2)}$ and
                        $\Phi.\reg(r_1,r_2) = \sealed{\sigma', c_1}, \sealed{\sigma', c_2}$
                        }\\
    \_                                           & $\failed$ & \totherwise \\
\begin{multline*}
  \srcxjmpres{c_1,c_2,\Phi} = \\
    \arraycolsep=1.4pt
    \begin{array}{l c >{\raggedright\arraybackslash}p{9.5cm}}
      \arraycolsep=0pt
    \begin{array}{rl}
      \Phi & \updReg{\pcreg}{c_1}\\
           & \updReg{\rdata}{c_2}
    &\phantom{mak} & $\nonExec{c_2}$ \srcalt{and $c_1 \neq \retptrc(\_)$ and $c_2 \neq \retptrd(\_)$}\\
      \arraycolsep=0pt
      \array[c]{rl}
      \Phi&\updReg{\pcreg}{c_\opc}\\
          &\updReg{\rstk}{c_\stk}\\
          &\updReg{\rdata}{0}\\
          &\update{\stk}{\stk'}\\
          &\update{\ms_\stk}{\ms_\stk \uplus \ms_\var{local}}
      \endarray
           & &
               $\arraycolsep=0pt\array{l}
        \src{(\opc, \ms_\var{local}) :: \stk'= \Phi.\stk \tand} \\
      \src{\var{retPtrMatchesOpc}(c_1,\opc)}\\
      \src{\var{retPtrForLocalStack}(c_2,\ms_\var{local})}\\
      \src{c_\stk = \var{reconstructStackPointer}(\reg,c_2) \tand}\\
      \src{c_\opc =  \var{tokenToPc}(c_1)} 
      \endarray$\\
      \failed &  & \totherwise
\right.
\end{multline*}
  \caption{An excerpt of the operational semantics of \srccm{} (some details omitted).}
  \label{fig:source-op-sem}

\begin{figure}
    \[
      \src{\var{sealedPairArgument}(\reg,r_1,r_2,c_1,c_2)} \mathit{ iff }
      \left\{ 
      \begin{array}{l}
        \src{\exists \sigma \ldotp (\reg(r_1) = \sealed{\sigma, c_1} \wedge \reg(r_2) = \sealed{\sigma, c_2}) \wedge{}}\\
        \src{r_1 \neq \rtmp{1} \wedge r_2 \neq \rtmp{2}}
      \end{array}
      \right.
    \]
    \[
      \src{ \opc, c_\opc = \var{setupOpc(((\_,\_),\baddr,\eaddr,\aaddr))}} \mathit{ iff }
      \left\{ 
      \begin{array}{l}
      \src{\opc = \aaddr + \calllen \wedge{}}\\
      \src{c_\opc = \retptrc(\baddr,\eaddr,\opc) \wedge{}}
      \end{array}
       \right.
    \]
    \begin{multline*}
      \src{\ms_{\stk,\var{local}},c_\var{local\_data},\ms_{\stk,\var{unused}},c_\stk = \var{splitStack}(\stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk}, \ms_\stk)} \mathit{\ iff} \\
      \left\{ 
      \begin{array}{l}
      \src{\baddr_\stk < \aaddr_\stk \leq \eaddr_\stk }\\
      \src{\ms_{\stk,\var{local}} = \ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42} }\\
      \src{\ms_{\stk,\var{unused}} = \ms_\stk|_{[\baddr_\stk,\aaddr_\stk-1]} }\\
      \src{c_\stk = \stkptr{\rw,\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1} }\\
      \src{c_\var{local\_data} = \retptrd(\aaddr_\stk,\eaddr_\stk) }\\
      \end{array}
      \right.
    \end{multline*}
    \[
      \src{\sigma = \var{getCallSeal}(\reg,\mem,\offpc,\offsigma)} \mathit{\ iff\ }
      \left\{
      \begin{array}{l} 
      \src{\reg(\pcreg) = ((\_,\_),\baddr,\eaddr,\aaddr) \wedge \baddr \leq  \aaddr+\offpc \leq \eaddr \wedge{}} \\
      \src{\mem(\aaddr+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\aaddr} \wedge{} \sigma_\baddr \leq \sigma \leq \sigma_\eaddr \wedge{}}\\
      \src{\sigma = \sigma_\aaddr + \offsigma }
      \end{array}
      \right.
    \]

@@

  \caption{The call condition in the operational semantics of \srccm{}.}
  \label{fig:source-op-sem}
\end{figure}
\begin{figure}
\[
    \src{\stkptr{\rw,\stkb,\eaddr_\stk,\aaddr_\stk} = \var{tokenToStkPtr}(\retptrd(\aaddr_\stk,\eaddr_\stk)}
\]
\[
    \src{((\rx,\normal),\baddr,\eaddr,\opc) = \var{tokenToPc}(\retptrc(\baddr,\eaddr,\opc)}
\]
% \[
%     \src{\stk', \ms_\var{local\_stk} = \var{popCallStackFrame} (\stk,c_1,c_2)}
%     \mathit{\ iff\ }
%       \left\{ \array{l}
%         \src{\stk = (\opc, \ms_\var{local\_stk}) :: \stk' \wedge{}} \\
%         \src{c_1 = \retptrc(\_,\_,\opc) \wedge{}} \\
%         \src{c_2 = \retptrd(\aaddr_\stk,\eaddr_\stk)\wedge{}} \\
%         \src{\dom(\ms_\var{local\_stk}) \subseteq [\aaddr_\stk,\eaddr_\stk]}
%         \endarray
%      \right.
% \]
\[
  \src{\var{retPtrMatchesOpc(\retptrc(\_,\_,\opc_c),\opc)}} \mathit{\ iff\ } \src{\opc_c = \opc}
\]
\[
  \src{\var{retPtrForMatchLocalStack(\retptrd(\aaddr_\stk,\eaddr_\stk),\ms_\var{local})}} \mathit{\ iff\ } \src{\dom(\ms_\var{local} \subseteq [\aaddr_\stk,\eaddr_\stk])}
\]

\[
    \src{c_\stk = \var{reconstructStackPointer}(\reg,c)}
    \mathit{\ iff\ }
      \left\{ \array{l}
        \src{c = \retptrd(\aaddr_\stk,\eaddr_\stk) \wedge{}} \\
        \src{\reg(\rstk) = \stkptr{\rw, \stkb, \aaddr_\stk-1,\_} \wedge{}} \\
        \src{\stkb \leq \addr_\stk \wedge{}} \\
        \src{c_\stk = \var{tokenToStkPtr(c)} }
        \endarray
     \right.
\]

% \begin{multline*}
%     \src{(\mem,\reg',\stk',\ms_\stk') = \var{returnCondition}(c_1,c_2,(\mem,\reg,\stk,\ms_\stk))}\mathit{\ iff\ }\\
%       \quad\quad\left\{ 
%       \begin{array}{l}
%       \src{\stk', \ms_\var{local\_stk} = \var{popCallStackFrame}(\stk,c_1,c_2)}\\
%       \src{c_\stk = \var{reconstructStackPointer}(\reg,c_2)}\\
%       \src{c_\opc =  \var{tokenToPc}(c_1)} \\
%       \src{\ms_\stk' = \ms_\stk \uplus \ms_\var{local\_stk}} \\
%       \src{\reg' =\reg\updReg{\pcreg,\rstk}{c_\opc,c_\stk}\updReg{\rdata,\rtmp{1},\rtmp{2}}{0,0,0}}
%       \end{array}\right.
%   \end{multline*}
